FAIR fission track analysis with geochron@home

### 2025-10-04 geochron@home/manuscript/paper.tex
Incorporated AC's suggestions

The combined data plots of Figures~\ref{fig:AvP}b and c contain two
sets of counts, with $\sum N_{1} = 686$ and $\sum N_{2} = 679$, and a
weighted mean track density ratio $\hat{\rho}_{1}/\hat{\rho}_{1} =
0.94$ with relative standard error 0.013. PV counted 600 tracks in
those common areas $A_0$, of which 549 were also counted by
AC. Conversely, AC counted 622 tracks of which 549 were also counted
by PV. The ratio of the two analysts' track density estimates based
just on the common area is therefore 600/622 = 0.965 with relative
standard error 0.018, calculated from Equation~\ref{eq:NO}, with $N_1
= 686$, $N_2 = 679$ and $N_0 = 549$.  This is close to the weighted
mean ratio of 0.94 (Figure~\ref{fig:AvP}c), and is slightly nearer to
1. It indicates that PV under-counts the Mount Dromedary apatite by
3.5\% relative to AC.  The existence of `observer bias' is well
documented \citep{tamer2025}.  It is one of the reasons why fission
track analysis is often done relative to age standards: observer bias
does not have to be a problem provided that it is consistent between
grains, and between samples.\medskip

@@

The annotations in the current tutorial pages were made by an
experienced fission track analyst (Andrew Carter). This basic tutorial
provides a quick and easy mechanism to train novice users in the art
of fission track analysis. The tutorial pages are in their infancy
and offer only a limited degree of interactivity. Users can click on
features to read the annotations. The tutorial will grow with input
from experienced analysts and future plans include the addition of
fully interactive `quizzes' (Section~\ref{sec:conclusions}). The
limitations of the current tutorials are apparent in the results of
the crowd-sourcing experiment described in the next section of this
paper.

@@

The fission track method has always been a test bed for new
geochronological developments.  Because fission track data are
imprecise, the fission track community has sollicited the help of
statisticians and mathematicians to develop its analytical protocols.
Other geochronological communities are still catching up with concepts
and tools such as overdispersion, mixture modelling and radial plots,
which have been commonplace in fission track analysis for decades
\citep{vermeesch2019b}. In a similar vein, the subjective nature of
fission track identification has prompted the fission track community
to organise inter-laboratory comparisons and round-robin studies long
before other geochronological communities
\citep{miller1985,tamer2025}.\medskip

>>>>

The combined data plots of Figures~\ref{fig:AvP}b and c contain two
sets of counts for 25 grains, with $\sum N_{1} = 686$ (PV) and $\sum
N_{2} = 679$ (AC), and a weighted mean track density ratio
$\hat{\rho}_{1}/\hat{\rho}_{1} = 0.94$ with relative standard error
0.013. PV counted 600 tracks in those common areas $A_0$, of which 549
were also counted by AC. Conversely, AC counted 622 tracks of which
549 were also counted by PV. The ratio of the two analysts' track
density estimates based just on the common area is therefore 600/622 =
0.965 with relative standard error 0.018, calculated from
Equation~\ref{eq:NO}, with $N_1 = 686$, $N_2 = 679$ and $N_0 = 549$.
This is close to the weighted mean ratio of 0.94
(Figure~\ref{fig:AvP}c), and is slightly nearer to 1. It indicates
that PV under-counts the Mount Dromedary apatite by 3.5\% relative to
AC.  The existence of `observer bias' is well documented
\citep{tamer2025}.  It is one of the reasons why fission track
analysis is often done relative to age standards: observer bias does
not have to be a problem provided that it is consistent between
grains, and between samples.\medskip

@@

Some exemplar annotations in the current tutorial pages were made by
an experienced fission track analyst (Andrew Carter). This basic
tutorial provides a quick and easy mechanism to train novice users in
the art of fission track analysis. The tutorial pages are in their
infancy and offer only a limited degree of interactivity. Users can
click on features to read the annotations. A more detailed,
comprehensive tutorial will grow with the wider input from experienced
analysts and future plans include the addition of fully interactive
`quizzes' (Section~\ref{sec:conclusions}). The limitations of the
current tutorials are apparent in the results of the crowd-sourcing
experiment described in the next section of this paper.

@@

The fission track method has always been a test bed for new
geochronological developments.  Because fission track data are
imprecise, the fission track community has solicited the help of
statisticians and mathematicians to develop its analytical protocols.
Other geochronological communities are still catching up with concepts
and tools such as overdispersion, mixture modelling and radial plots,
which have been commonplace in fission track analysis for decades
\citep{vermeesch2019b}. In a similar vein, the subjective nature of
fission track identification has prompted the fission track community
to organise inter-laboratory comparisons and round-robin studies long
before other geochronological communities
\citep{miller1985,tamer2025}.\medskip
